Whittaker Models for Real Groups

### 2024-06-26 whittaker.tex
Small edits

Suppose $\pi$ is a discrete series representation of~$G(\R)$, and let $\lambda\in\t^*$ be a representative of the Harish-Chandra parameter of $\pi$.
Define\footnote{AA 2024-VI-14: should perhaps think about consistency in notation, and decide if we fix a large chamber from the start and take the HC parameter to live there, or add a few words about independence of choices...}
\begin{equation} \label{semisimple_orbit_HC} \mathcal{O}_{\pi}=G(\R)\cdot\lambda.\end{equation}

@@

The complexified Lie algebra $\g$ is the space of $2x2$ complex matrices of trace $0$. Let $\s(\C)=\{t_z\mid z\in\C\}\subset \g$ where
$$
t_z=\begin{pmatrix}0&z\\-z&0
\end{pmatrix}
$$
Then $\s(\R)=\{t_x\mid x\in\R\}$.

@@

The Killing form satisfies
$$
\kappa(t_x,t_y)=-8xy.
$$
Define $H_{\lambda}\in \s(\C)$ so that $\kappa(H_\lambda,t_x)=\lambda(t_x)$. Then $H_\lambda=t_w$ for some $w$; so
$$
\kappa(H_{\lambda_z},t_x)=\lambda_z(t_x)=zx,\quad \kappa(t_w,t_x)=-8xw
$$
gives
$$
H_{\lambda_z}=t_{-\frac z8}.
$$
In particular
$$
H_{\pm \rho}=H_{\lambda_{\mp i}}=\begin{pmatrix}0&-\frac {\mp i}8\\\frac {\mp i}8&0
\end{pmatrix}
$$
Let $\pi(\lambda_{ik})$ be the discrete series representation with Harish-Chandra parameter $\lambda_{ik}$ $(k\in \Z_{\ne 0})$, so
$$
H_{\pi(\lambda_{ik})}=H_{\lambda_{ik}}=
\begin{pmatrix}0&-\frac{ik}8\\\frac{ik}8&0
\end{pmatrix}
$$
For an $SL(2)$-triple we can take $\{X_\alpha,X_{-\alpha},t_{-i}\}$ where
$$
X_\alpha=\frac12\begin{pmatrix}1&i\\i&-1
\end{pmatrix}, \quad X_{-\alpha}=\overline{X_\alpha}, \quad t_{-i}=\begin{pmatrix}0&-i\\i&0
\end{pmatrix}
$$
That is
$$
[t_z,X_\alpha]=\alpha(t_z)X_\alpha=\lambda_{2i}(t_z)X_\alpha=2izX_\alpha
$$
In particular
$$
[t_{-i},X_\alpha]=2X_\alpha
$$

Let's compute $\AV(\pi(\lambda_{ik})$ using  \eqref{e:Fpi}.
Suppose $k>0$. Then $\langle \lambda_{ik},\ch\alpha\rangle=k$, so $\Delta^+=\{\alpha\}$, and
$F_{\pi_{ik}}=X_{-\alpha}$. 
By \eqref{e:Fpi} $\AV(\pi_{ik})$ is the $K$-orbit of $X_{-\alpha}^\ast$.
Thus in general
$$
\AV(\pi_{ik})=K\cdot X^*_{-\sgn(k)\alpha}
$$
Note: this only uses \eqref{e:Fpi}, which comes directly from \cite{AV1}.

Conjugating $H_{\pi(\lambda_{ik})}$ by $\mathrm{diag}(x,\frac 1x)$ takes it to
$\begin{pmatrix}0&-x^2\frac{ik}8\\\frac{ik}{x^2}&0
\end{pmatrix}$,
and taking the limit we see
$$
\WF(\pi(\lambda_{ik}))=\AC(G(\R)\cdot H_{\pi(\lambda_{ik}}))=
G(\R)\cdot \begin{pmatrix}0&-\sgn(k)i\\0&0
\end{pmatrix}
$$
Note: this only uses  \cite{HarrisHeOlafsson}.

Note that the $G(\R)$ orbit of $\begin{pmatrix}0&i\\0&0
\end{pmatrix}$ contains $\begin{pmatrix}0&i\R^+\\0&0
\end{pmatrix}$,
so in particular $G(\R)\cdot \begin{pmatrix}0&ix\\0&0
\end{pmatrix}=G(\R)\cdot \begin{pmatrix}0&iy\\0&0
\end{pmatrix}$ if and only if $\sgn(x)=\sgn(y)$.
In particular
$$
\WF(\pi_{ik})=\WF(\pi_{i\ell})\Leftrightarrow \sgn(k)=\sgn(\ell)
$$

Finally let's compute $\Wh(\pi_{\lambda_{ik}})$ in two ways.
First of all by Lemma \ref{lem:matumoto}, which comes directly from \cite[Theorem A]{matumoto}
$$
\Wh(\pi_{\lambda_{ik}})=\WF(\pi_{\lambda_{ik}})=G(\R)\cdot \begin{pmatrix}0&-\sgn(k)i\\0&0
\end{pmatrix}
$$

@@

$$
\O_{\pi_{\lambda_{ik}}}=G(\R)\cdot H_{\lambda_{ik}}=G(\R)\cdot \begin{pmatrix}0&-\frac{ik}8\\\frac{ik}8&0
\end{pmatrix}
$$

Note that
$$
\K
\begin{pmatrix}0&\pm i\\0&0
\end{pmatrix}=\{\begin{pmatrix}0&\pm i\\z&0
\end{pmatrix}\mid z\in\C\}
$$
so, with $\epsilon=\pm 1$, 
$$
H_{\lambda_{ik}}\in \K\begin{pmatrix}0&\epsilon i\\0&0
\end{pmatrix}\Leftrightarrow \epsilon=\sgn(k).
$$

By Proposition \ref{JeffTasho_criterion}
we conclude (again) that
$$
\Wh(\pi_{\lambda_{ik}})=G(\R)\cdot\begin{pmatrix}0&-\sgn(k)i\\0&0
\end{pmatrix}
$$

>>>>

Suppose $\pi$ is a discrete series representation of~$G(\R)$, and let $\lambda\in\t^*$ be a representative of the Harish-Chandra parameter of $\pi$.
Define
\begin{equation} \label{semisimple_orbit_HC} \mathcal{O}_{\pi}=G(\R)\cdot\lambda.\end{equation}
This depends only on~$\pi$, and not on the choice of representative~$\lambda$.

@@

The complexified Lie algebra $\g$ is the space of $2\times 2$ complex matrices of trace~$0$. Let $\s(\C)=\{t_z\mid z\in\C\}\subset \g$ where
$$
t_z=\begin{pmatrix}0&z\\-z&0
\end{pmatrix}.
$$
Then $\s(\R)=\{t_x\mid x\in\R\}$.

@@

The Killing form satisfies
$$
\kappa(t_x,t_y)=-8xy.
$$
Define $H_{\lambda}\in \s(\C)$ so that $\kappa(H_\lambda,t_x)=\lambda(t_x)$. Then $H_\lambda=t_w$ for some $w$; so
$$
\kappa(H_{\lambda_z},t_x)=\lambda_z(t_x)=zx,\quad \kappa(t_w,t_x)=-8xw
$$
gives
$$
H_{\lambda_z}=t_{-\frac z8}.
$$
In particular
$$
H_{\pm \rho}=H_{\lambda_{\mp i}}=\begin{pmatrix}0&-\frac {\mp i}8\\\frac {\mp i}8&0
\end{pmatrix}
$$
Let $\pi(\lambda_{ik})$ be the discrete series representation with Harish-Chandra parameter $\lambda_{ik}$ $(k\in \Z_{\ne 0})$, so
$$
H_{\pi(\lambda_{ik})}=H_{\lambda_{ik}}=
\begin{pmatrix}0&-\frac{ik}8\\\frac{ik}8&0
\end{pmatrix}.
$$
For an $SL(2)$-triple we can take $\{X_\alpha,X_{-\alpha},t_{-i}\}$ where
$$
X_\alpha=\frac12\begin{pmatrix}1&i\\i&-1
\end{pmatrix}, \quad X_{-\alpha}=\overline{X_\alpha}, \quad t_{-i}=\begin{pmatrix}0&-i\\i&0
\end{pmatrix}.
$$
That is
$$
[t_z,X_\alpha]=\alpha(t_z)X_\alpha=\lambda_{2i}(t_z)X_\alpha=2izX_\alpha.
$$
In particular
$$
[t_{-i},X_\alpha]=2X_\alpha.
$$

Let's compute $\AV(\pi(\lambda_{ik}))$ using  \eqref{e:Fpi}.
Suppose $k>0$. Then $\langle \lambda_{ik},\ch\alpha\rangle=k$, so $\Delta^+=\{\alpha\}$, and
$F_{\pi_{ik}}=X_{-\alpha}$. 
By \eqref{e:Fpi} $\AV(\pi_{ik})$ is the $K$-orbit of $X_{-\alpha}^\ast$.
Thus in general
$$
\AV(\pi_{ik})=K\cdot X^*_{-\sgn(k)\alpha}
$$
Note: this only uses \eqref{e:Fpi}, which comes directly from \cite{AV1}.

Conjugating $H_{\pi(\lambda_{ik})}$ by $\mathrm{diag}(x,\frac 1x)$ takes it to
$\begin{pmatrix}0&-x^2\frac{ik}8\\\frac{ik}{x^2}&0
\end{pmatrix}$,
and taking the limit we see
$$
\WF(\pi(\lambda_{ik}))=\AC(G(\R)\cdot H_{\pi(\lambda_{ik})})=
G(\R)\cdot \begin{pmatrix}0&-\sgn(k)i\\0&0
\end{pmatrix}
$$
Note: this only uses  \cite{HarrisHeOlafsson}.

Note that the $G(\R)$ orbit of $\begin{pmatrix}0&i\\0&0
\end{pmatrix}$ contains $\begin{pmatrix}0&i\R^+\\0&0
\end{pmatrix}$,
so in particular $G(\R)\cdot \begin{pmatrix}0&ix\\0&0
\end{pmatrix}=G(\R)\cdot \begin{pmatrix}0&iy\\0&0
\end{pmatrix}$ if and only if $\sgn(x)=\sgn(y)$.
In particular
$$
\WF(\pi_{ik})=\WF(\pi_{i\ell})\Leftrightarrow \sgn(k)=\sgn(\ell).
$$

Finally let's compute $\Wh(\pi_{\lambda_{ik}})$ in two ways.
First of all by Lemma \ref{lem:matumoto}, which comes directly from \cite[Theorem A]{matumoto}:
$$
\Wh(\pi_{\lambda_{ik}})=\WF(\pi_{\lambda_{ik}})=G(\R)\cdot \begin{pmatrix}0&-\sgn(k)i\\0&0
\end{pmatrix}.
$$

@@

$$
\O_{\pi_{\lambda_{ik}}}=G(\R)\cdot H_{\lambda_{ik}}=G(\R)\cdot \begin{pmatrix}0&-\frac{ik}8\\\frac{ik}8&0
\end{pmatrix}.
$$
On the other hand
$$
\K
\begin{pmatrix}0&\pm i\\0&0
\end{pmatrix}=\{\begin{pmatrix}0&\pm i\\z&0
\end{pmatrix}\mid z\in\C\}
$$
so, with $\epsilon=\pm 1$, 
$$
H_{\lambda_{ik}}\in \K\begin{pmatrix}0&\epsilon i\\0&0
\end{pmatrix}\Leftrightarrow \epsilon=\sgn(k).
$$

By Proposition \ref{JeffTasho_criterion}
we conclude (again) that
$$
\Wh(\pi_{\lambda_{ik}})=G(\R)\cdot\begin{pmatrix}0&-\sgn(k)i\\0&0
\end{pmatrix}.
$$
Alexandre: the proof of Proposition~\ref{JeffTasho_criterion} uses Lemma~\ref{lem:matumoto}.
